\mysection{Executable files patching}
\label{patching}

\subsection{x86 code}
\label{x86_patching}

Frequent patching tasks are:

\myindex{x86!\Instructions!NOP}
\begin{itemize}

\item 
One of the most frequent jobs is to disable some instruction.
It is often done by filling it using byte 
\TT{0x90} (\ac{NOP}).

\item Conditional jumps, which have an opcode like \TT{74 xx} (\JZ), 
can be filled with two \ac{NOP}s.

It is also possible to disable a conditional jump by writing 0 at the second byte (\emph{jump offset}).

\myindex{x86!\Instructions!JMP}
\item 
Another frequent job is to make a conditional jump to always trigger: 
this can be done by writing \TT{0xEB} 
instead of the opcode, which stands for \JMP.

\myindex{x86!\Instructions!RET}
\myindex{stdcall}
\item A function's execution can be disabled by writing \RETN (0xC3) at its beginning.
This is true for all functions excluding \TT{stdcall} (\myref{sec:stdcall}).
While patching \TT{stdcall} functions, one has to determine the number of arguments (for example, 
by finding \RETN in this function), 
and use \RETN with a 16-bit argument (0xC2).

\myindex{x86!\Instructions!MOV}
\myindex{x86!\Instructions!XOR}
\myindex{x86!\Instructions!INC}
\item Sometimes, a disabled functions has to return 0 or 1.
This can be done by \TT{MOV EAX, 0} or \TT{MOV EAX, 1}, 
but it's slightly verbose.\\
A better way is \TT{XOR EAX, EAX} (2 bytes \TT{0x31 0xC0}) or \TT{XOR EAX, EAX / INC EAX} (3 bytes \TT{0x31 0xC0 0x40}).

\end{itemize}

A software may be protected against modifications.

This protection is often done by reading the executable code and calculating a checksum.
Therefore, 
the code must be read before protection is triggered.

This can be determined by setting a breakpoint on reading memory.

\myindex{tracer}
\tracer has the BPM option for this.

PE executable file relocs (\myref{subsec:relocs}) 
must not to be touched while patching, 
because the Windows loader may overwrite your new code.
\myindex{Hiew}
(They are grayed in Hiew, for example:
\figref{fig:scanf_ex3_hiew_1}).

As a last resort, it is possible to write jumps that circumvent the relocs, 
or you will have to edit the relocs table.

