% To translators: don't bother to translate this... english-only version.

\begin{center}
\LARGE{} This is my own bulletin board \normalsize{}
\end{center}

\textbf{Please donate.}
... so I could continue working on my writings...\\
Bitcoin: \textbf{1LLa7aqQbRmCYbccnCNwaLgxK9jFPTqPw}
If you want to get into list of donors, please drop me email about that.

Sorry, no PayPal, citizens of Ukraine can't receive payments.
Maybe someone can help me with that?
Collect PayPal donations and then transfer them to me via Bitcoin?

-> \EMAIL{}

\myhrule{}

I would be happy to hear from anyone interested in my consulting services.
The area of my expertise is programming in general, reverse engineering, maybe technical writing...

\myhrule{}

\textbf{This book is probably outdated already}.
(Unless it was just downloaded from \url{https://beginners.re/}.)

The book is \href{https://github.com/DennisYurichev/RE-for-beginners/commits/master}{changing too often},
content being added, bugs are (hopefully) being fixed.
The latest version is always at \url{https://beginners.re/}.

This PDF you currently reading was compiled at \today{}.

\myhrule{}

My dear readers! From time to time, I have questions, I don't know who (or where) to ask.
Or I'm just lazy...
Can you please help me?

\myhrule{}

How do you install VMware Remote Console 10.0.4 on Ubuntu 19? It just suddenly exits during installation.
Is it known symptom?

Or what do you use to run VMware Workstation VMs on remote Ubuntu box?

\myhrule{}

What do you use on Linux in place of Adobe Acrobat Pro?
To edit contents, add bookmarks, notes, highlight text...

\myhrule{}

What does tilde means in versions number in .dsc files, which describing Ubuntu packages?
Some kind of wildcard?
And what does \verb|>>| means?
\emph{Pipe} is just \emph{OR}?
For example:

\begin{lstlisting}
Build-Depends: cython-dbg | python-pyrex, ca-certificates, debhelper (>> 8.1.0~), python (>= 2.6.6-3), python-all-dev (>= 2.6.6-3), python-all-dbg (>= 2.6.6-3), python-configobj (>= 4.7.2+ds-2), python-docutils, python-paramiko, python-pycurl-dbg, python-subunit, python-testtools (>= 0.9.5~)
\end{lstlisting}

\myhrule{}

What do you use on unrooted Android device as a SSH server, that can write to external micro-SD card?
I was happy with SSHDroid, but now it's outdated...
For instance, when Total Commander writes to micro-SD, a dialog box appears, whether to allow this or not.
You allow it, and it can write to any folder on flash.
This is not a case with SSH-servers that I saw.

\myhrule{}

A win32 process A is running.
Process B is attaching to it as a debugger, or opens it using OpenProcess().
ReadProcessMemory() works OK, but fails if it tries to read uncommited memory pages of process A.

The problem: how to force the Windows Memory Manager to commit a page in process A from userland of process B?
I can inject a read instruction into process A, run it, and the page would be committed, but this is not the solution.

\myhrule{}

If you know something, please help me: \EMAIL{}

